---
title: "Practical Machine Learning Course Project"
author: "Jijo Jose"
date: "14 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading the data

```{r reading data}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

# Preprocessing

## Removing columns with 70% or more NA values 

 Cloumns with 70% of NA values are not useful for the creation of the model. So, it is better to remove those columns. 
 
```{r removing columns with 70% NA values, cache=TRUE}
#importing the required library
library(caret)

#Selecting columns with more than 70% of non null values for analysis
selected_columns <- colSums(!is.na(training))/nrow(training) >= 0.6
trainingData <- training[,selected_columns]
```

## Removing columns with near zero variance

Variables without any significant variance are not useful for the creation of the model. So, we are removing those.

```{r Removing columns with approximately zero variance}
library(caret)
#Selecting columns with near zero variance
nzv <- nearZeroVar(trainingData)
trainingData <- trainingData[,-nzv]
```

##Removing columns of time series and name

Columns of time series are an indicator of when the event happened. Since we want to predict the activity by the data generated from sensors, adding those variables won't give the required analysis results for us. So, we are removing it. We are also removing the name variable.

```{r Removal of time series}
# Reoving time series data and name
trainingData <- trainingData[,-c(1:6)]
```

##Saving the column names used for modelling

The column names selected for the creation of the model are required for prediction. So, we are saving them.

```{r saving column names}
colSelected <- colnames(trainingData)
```

## Dividing the training data into trainingset and testingset

Inorder to evaluate the model, we need a testing set. So, we will divide the training data into trainingset and testingset.

```{r trainingset and testing set}
# Training set
tr <- createDataPartition(trainingData$classe,p=0.7,list = FALSE )
trainingSet <- trainingData[tr,] 
testingSet <- trainingData[-tr,]

```

# Creating  the model

Now, let us create the model. We will be using the Random Forest algorithm from the caret package for model creation.

```{r model creation, cache=TRUE}
library(caret)

#Creating model
model <- train(classe~., data = trainingSet, method = "rf", trControl=trainControl(method = "cv"))

# The model
model
```

# Model Evaluation

Now, let us evaluate the performance of the model using the testingset.

```{r model evaluation, cache=TRUE}
#Predicting with testing set
pr <- predict(model, newdata = testingSet)

#Accuracy on testing set
confusionMatrix(pr,testingSet$classe)
```

The model is giving an accuracy greater than 99%. So, we can conclude that the model can predict the activities pretty well.

# Predicting for the dataset "testing"
 
Now, let's predict for the 20 test cases. 

```{r predicting with testing set}
#Selecting the required columns
testing <- testing[,colSelected[1:52]]

#Predicting for the testing set
prt <- predict(model, newdata = testing)

#The predictions
prt
```

So, these are the final predictions for the testing set. 